Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-05-06 WSSSPE2_report/WSSSPE2_review.tex
Reformat. No content change.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Lightning Talks} \label{sec:lightning}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{comment}
\note{Lead: Choi. Volunteers welcome.
\href{http://wssspe.researchcomputing.org.uk/wssspe2/lightning-talks/}{Slides.}}
\end{comment}

\begin{comment}
Lighting talks, breakout groups, working groups are all different. Lightning
talks were new in WSSSPE2. They were offered to all paper contributors, and some
other attendees, those who got travel awards. Breakout groups in WSSSPE1 were
just discussions about topics; in WSSSPE2, the breakout groups were supposed to
come up with plans for actions in specific areas. The idea of working groups is
that they will come out of breakout groups as a way for the groups to actually
do the actions over time, not at the workshops.

@@

%\item Christopher Lenhardt: Open Science for Synthesis (OSS): Filling the Gap in
%Early Career Training

@@

\subsection{Robert Downs: Community Recommendations for Improving Sustainable
Scientific Software Practices~\cite{Downs_poster}}
Robert Downs, of the Columbia University Center for International Earth
Science Information Network (CIESIN), 
%presented Community Recommendations for
%Improving Sustainable Scientific Software Practices (2014), and 
described a
focus group study conducted with %W. Christopher Lenhardt, Erin Robinson, Ethan
% Davis, and Nicholas Weber of 
the Science Software Cluster (SSC) of the
Federation of Earth Science Information Partners (ESIP). For the study, almost
300 attendees of the 2014 Summer ESIP Meeting were invited to participate in
simultaneous roundtable discussions on sustainability of science software. Over
two-thirds of the roundtable focus groups responded to a semi-structured survey
that contained three sets of questions eliciting recommendations for near-term
actions of the community to improve sustainable software practices. Initial
analysis of the
participants' responses to the questionnaire revealed several suggestions, which
included improving community engagement and collaborative activities, increasing
understanding and awareness, and creating incentives to motivate sustainable
science software practices.
%Based on the recommendations
%offered by the focus group respondents for near-term actions, 
The ESIP SSC plans to engage the community %to achieve progress in the near future 
in the recommended activities for improving sustainable scientific
software practices.

@@

\subsection{Jakob Blomer: The Need for a Versioned Data Analysis Environment~\cite{Blomer_poster}}
%
Large-scale scientific endeavors, such as the discovery of the Higgs
boson at the Large Hadron Collider (LHC), often rely on complex software
stacks.  Maintaining thousands of dependencies of software
libraries and operating system versions has shown that despite
source code availability, the setup and the validation of a minimal
usable analysis environment can easily become prohibitively expensive.
%There is a substantial gap between merely having access to open source
%software and the ability to create a data analysis runtime environment.
 In high-energy physics, a
special purpose, open-source, versioning and snapshotting file system
(CernVM-FS) that is used to capture and distribute entire software stacks,
providing instant access to ready-to-run data analysis environments,
proved to be very useful.
% for the global scientific community as well as for
%interested citizens, for instance through the CERN Open Data Portal
%(http://opendata.cern.ch). \todo{revise}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Lightning Talks} \label{sec:lightning}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{comment}
\note{Lead: Choi. Volunteers welcome.
\href{http://wssspe.researchcomputing.org.uk/wssspe2/lightning-talks/}{Slides.}}

Lighting talks, breakout groups, working groups are all different. Lightning
talks were new in WSSSPE2. They were offered to all paper contributors, and some
other attendees, those who got travel awards. Breakout groups in WSSSPE1 were
just discussions about topics; in WSSSPE2, the breakout groups were supposed to
come up with plans for actions in specific areas. The idea of working groups is
that they will come out of breakout groups as a way for the groups to actually
do the actions over time, not at the workshops.

@@

%\item Christopher Lenhardt: Open Science for Synthesis (OSS): Filling the Gapin %Early Career Training

@@

\subsection{Robert Downs: Community Recommendations for Improving Sustainable
Scientific Software Practices~\cite{Downs_poster}}
Robert Downs, of the Columbia University Center for International Earth Science
Information Network (CIESIN),
%presented Community Recommendations for
%Improving Sustainable Scientific Software Practices (2014), and 
described a
focus group study conducted with %W. Christopher Lenhardt, Erin Robinson, Ethan
% Davis, and Nicholas Weber of 
the Science Software Cluster (SSC) of the Federation of Earth Science
Information Partners (ESIP). For the study, almost 300 attendees of the 2014
Summer ESIP Meeting were invited to participate in simultaneous roundtable
discussions on sustainability of science software. Over two-thirds of the
roundtable focus groups responded to a semi-structured survey that contained
three sets of questions eliciting recommendations for near-term actions of the
community to improve sustainable software practices. Initial analysis of the
participants' responses to the questionnaire revealed several suggestions, which
included improving community engagement and collaborative activities, increasing
understanding and awareness, and creating incentives to motivate sustainable
science software practices. The ESIP SSC plans to engage the community in the
recommended activities for improving sustainable scientific software practices.

@@

\subsection{Jakob Blomer: The Need for a Versioned Data Analysis Environment~\cite{Blomer_poster}}
%
Large-scale scientific endeavors, such as the discovery of the Higgs boson at
the Large Hadron Collider (LHC), often rely on complex software stacks.
Maintaining thousands of dependencies of software libraries and operating system
versions has shown that despite source code availability, the setup and the
validation of a minimal usable analysis environment can easily become
prohibitively expensive. In high-energy physics, a special purpose, open-source,
versioning and snapshotting file system (CernVM-FS) that is used to capture and
distribute entire software stacks, providing instant access to ready-to-run data
analysis environments, proved to be very useful.
% for the global scientific community as well as for
%interested citizens, for instance through the CERN Open Data Portal
%(http://opendata.cern.ch). \todo{revise}
